---
title: "hw02 - Shot Charts"
author: "Hanmaro Song"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = '../images/')
```

```{r}
library(dplyr)
library(ggplot2)
shots_data = read.csv('../data/shots-data.csv', stringsAsFactors = FALSE)

```


```{r}
# Question 5.1
total_shots <- shots_data %>% 
  group_by(name) %>% 
  summarise(total = n(), na.rm=TRUE) %>% 
  select(name, total) %>% 
  arrange(desc(total))
total_shots

# Question 5.2
effective <- shots_data %>% 
  group_by(name) %>% 
  filter(shot_made_flag == 'made shot') %>% 
  summarise(made = n(), na.rm=TRUE) %>% 
  select(name, made) %>% 
  inner_join(total_shots) %>% 
  select(name, total, made) %>%
  mutate(perc_made = made / total) %>% 
  arrange(desc(perc_made))
effective


two_point <- shots_data %>% 
  filter(shot_type == '2PT Field Goal' & shot_made_flag == 'made shot') %>% 
  group_by(name) %>% 
  summarise(made = n(), na.rm=TRUE) %>% 
  select(name, made) %>%
  inner_join(total_shots,two_point, by = 'name') %>% 
  select(name, total, made) %>%
  mutate(perc_made=made/total) %>% 
  arrange(desc(perc_made))
two_point

three_point <- shots_data %>% 
  filter(shot_type == '3PT Field Goal' & shot_made_flag == 'made shot') %>% 
  group_by(name) %>% 
  summarise(made = n(), na.rm=TRUE) %>% 
  select(name, made) %>%
  inner_join(total_shots,three_point, by = 'name') %>% 
  select(name, total, made) %>%
  mutate(perc_made=made/total) %>% 
  arrange(desc(perc_made))
three_point



```


```{r}
# Question 6.1

made_shots <- shots_data %>% 
  group_by(shot_distance) %>% 
  summarise(total = n(), made = sum(shot_made_flag == 'made shot'), na.rm = TRUE) %>% 
  select(shot_distance,total,made) %>% 
  mutate(made_shot_prop = made / total) %>% 
  select(shot_distance,made_shot_prop)
made_shots

```


```{r plotchart_distance_to_made_shot}
# Question 6.2
ggplot(made_shots) +
  geom_point(aes(shot_distance, made_shot_prop))

```
1. When the distance is at 0, it has the highest chance to score. Then distance at 1 and 2 follows. From about 3 or 4 to 33, it has around slightly less than 50%. After that there is almost no chance at all to score except at 56 in which case there was one attempt that resulted in made shot.

2. By the given graph, yes. Although some distance range (such as from 3 to 33?) has similar probability.

3. From the distance at around 33ft, the probability can be considered as null.

4. At 0, 1, 2(or upto 3), around 12, 16, 17, 21, 30 have 50% or more. 

```{r chart_player_shots_per_min}
rect1 = data.frame(xmin=0, xmax = 12, ymin = 0, ymax = 60)
rect2 = data.frame(xmin=24, xmax = 36, ymin = 0, ymax = 60)

dat = select(shots_data, name, period, minutes_remaining, shot_made_flag)
dat = mutate(dat, time = period*12-minutes_remaining) %>% select(name, time)
dat = group_by(dat, name) %>% count(time)
dat = select(dat, name, time, shots = n)

chart_player_shots_per_min = ggplot(data = dat, aes(x=time, y=shots)) +
  geom_rect(data = rect1, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill='gray', alpha=0.4, inherit.aes = FALSE) +
  geom_rect(data = rect2, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill='gray', alpha=0.4, inherit.aes = FALSE) +
  geom_path(aes(x = time, y = shots),colour='blue',alpha=0.5) +
  geom_point(aes(x = time, y = shots),colour='blue',alpha=0.5) +
  facet_wrap(~ name) +
  ylim(0, 60) +
  scale_x_continuous( name = "minute", breaks=c(1,12,24,36,48)) +
  theme_minimal() +
  ylab("total number of shots")
chart_player_shots_per_min

```


































