

```{r, echo = FALSE, message = FALSE}

library(dieroller)
```

## Die Rolling

The package `"dieroller"` is a minimal implementation for simulating rolling
a die a given number of times, and to visualize the relative frequencies
in such a series of rolls.


## Creating a `die`

The first step is to create a `"die"` object with the function `die()`:

```{r}
die1 <- die()
die1
```

By default, `die()` creates a _fair_ die with standard sides 1 through 6,
and with probability values for each side of 0.166667.

You can change the defatuls with arguments `sides` and `prob`:

```{r}
loaded <- die(sides = c('a', 'b', 'c', 'd', 'e', 'f'), prob = c(0.1, 0.1, 0.2, 0.2, 0.1, 0.3))
loaded
```


## Rolling a die

Once you have defined a `"die"` object, you can create `roll()` object to
roll a die a given number of `times`. The output will be an object of class 
`"roll"`, which will contain the vector of `rolls`:

```{r}
roll100 <- roll(die1, times = 100)
roll100
```

The `summary()` function gives you a nice table of frequencies:
```{r}
summary(roll100)
```


## Plot of relative frequencies

You can use the function `plot()` to create a graohic with the relative 
frequencies of a series of coin tosses:

```{r, fig.show='hold'}
plot(roll100)
```
