---
title: "lab01"
author: "Hanmaro Song"
date: "January 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
phone <- 110
transportation <- 650
groceries <- 500
gym <- 70
rent <- 1000
other <- 800

expenses <- sort(c(phone, transportation, groceries, gym, rent, other), TRUE)
names <- c('phone', 'transportation', 'groceries', 'gym', 'rent', 'other')
chart <- barplot(expenses, names.arg = names, cex.names = 0.9, ylim = c(0, 1100))
text(chart, expenses, labels = expenses, pos = 3, col = 'red')
```



```{r}
# Pythagoras Formula
a = 3
b = 4
c = sqrt(a^2 + b^2)
print(c)


# Binomial Formula
n = 5
k = 2
p = 0.5

combinations = factorial(n) / (factorial(k) * factorial(n-k))
r1 = combinations * p^k * (1 - p)^(n-k)
r2 = choose(5, 2) * p^k * (1 - p)^(n-k)
print(r1 == r2)

roll_ten = choose(10, 3) * (1/6)^3 * (1 - (1/6))^(7)
print(roll_ten)

r3 = dbinom(2, 5, 0.5)
r4 = dbinom(3, 10, (1/6))
print(r3)
print(r4)

biased = dbinom(2, 5, .35)
print(biased)

more_than_four = dbinom(4, 5, .35) + dbinom(5, 5, 0.35)
print(more_than_four)

###########################
# Questions after installing the libraries
x = 2
r1 = 3*x^2 + 4*x + 8
x <- -3:3
r2 = 3*x^2 + 4*x + 8
print(r1)
print(r2)

# One can use help.search() and type any keyword that (s)he wants to find. ex) help.search('operator') 
# Files tab is good when navigating through files and load up a file instead of minimizing the RStudio
# When I have no clue as to how to use a certain function (command) or look for a function with search
# History is good when I need to check what I've run in the past.
# Navigate the current display all the way back to home directory of RStudio where one can easily click a link to view wanted page.
# In environment tab, one can view all the variables declared. 
