---
title: "README"
author: "Hanmaro Song"
date: "April 23, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE)
```



## Overview

`"dieroller"` is a minimal [R](http://www.r-project.org/) package 
that provides functions to simulate rolling a die.

* `die()` creates a die object (of class `"die"`)
* `roll()` rolls a die object, producing a `"roll"` object.
* `plot()` method for a `"roll"` object to plot frequencies of sides of a die.
* `summary()` method for a `"roll"` object.


## Motivation

This package has been developed to illustrate some of the concepts
behind the creation of an R package.


## Installation

Install the development version from GitHub via the package `"devtools"`:

```r
# development version from GitHub:
#install.packages("devtools") 


# install "dieroller"
devtools::install_github("stat133-sp18/hw-stat133-hsong1101/hw5", build_vignettes = TRUE)
```


## Usage

```{r}
library(dieroller)

# default die1
die1 = die()
die

# default die2
die2 = die(sides = c('a', 'b', 'c', 'd', 'e', 'f'), prob = c(0.1, 0.2, 0.3, 0.2, 0.1, 0.1))
die2

# default roll
roll1 = roll(die1)
roll2 = roll(die2, times = 10)

# default roll print
roll1
roll2

# summary of roll
summary(roll2)

# barplot of roll
plot(roll2)

# addition of rolls
roll2 = roll2 + 100
summary(roll2)

```




































